Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-21 paper/paper.tex
Update section 6

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enables the algorithms to finish almost all the workflows. In three of five 
cases the SPSS algorithm performs better for tighter constraints: when the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. The reason for 
this behavior is that the large number of tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the tested workflows the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

@@

In this paper we addressed a new interesting and important problem of scheduling
and resource provisioning for scientific workflow ensembles on IaaS cloud
infrastructures. The problem is different from previous work on grid or utility
grid scheduling that cloud infrastructure gives more control over the resources,
so the resource provisioning plan can be adjusted to the application
requirements. Therefore the problem space becomes larger: it requires not
only to select the best scheduling of tasks to available resources, but also
select the best shape of resource provisioning plan. The formulation of a
problem as maximization of the number of prioritized workflows completed from
the ensemble requires also to take the decisions on admitting or rejecting workflows
based on their estimated resource demands. We believe that such formulation of
the bi-constrained problem is highly relevant since such
constraints are typically imposed on many real-world projects.

We have analyzed both static and dynamic scheduling approaches and
developed the SPDS, DPDS, WA-DPDS and SPSS algorithms realizing these
strategies. The algorithms were evaluated on ensembles of benchmark
workflows which represent typical real scientific applications. For the
purpose of evaluation we have developed a simulator that models the cloud
infrastructure and the tightly-coupled scheduler and provisioner modules of
workflow engine. The results of simulation studies indicate that the two algorithms (WA-DPDS and
SPSD) that take into account the information about the workflow structure and
estimation of task runtimes yield better results than the simple on-line
priority-based scheduling strategy with static (constant) resource provisioning
plan.

During our work, several interesting lessons were learned about the workflow
ensembles on cloud infrastructures. One result is that the simple strategy of
SPDS performs relatively well in terms of number of workflows completed and the
high resource utilization it can achieve. However, this method does not allow to
give any advance prediction about the result, so to see its performance the
whole ensemble needs to be simulated. On the other hand, the method may be
the only choice in the case no estimates about the task runtimes are available
in advance. It can be seen that the workflow-aware and static algorithms are
able to find better quality solutions in terms of the defined metrics: this the
result mainly of the admission algorithm implemented, which allow rejecting too
large workflows and thus saving resources for completing the less costly ones.

Another general observation is that although clouds provide such capabilities as
dynamic adding and removing of resources at runtime and high parallelization due
to illusion of unlimited resources, it turns out that is recommended to avoid
using these capabilities when possible, in order to achieve high resource
utilization and cost effectiveness. This observation comes from the fact that
starting and stopping more VMs incurs a cost which can be amortized by running
each started VM as long as possible before shutdown. This effect was confirmed
in the case of SPSS algorithm for Montage and CyberShake workflows with small
budgets. Similarly, high degree of parallelization (large number of nodes) leads
also to less efficient resource utilization and in the case of clouds the drop
in parallel efficiency means also the increase of cost. This result, clearly
visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
related trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

Another lesson learned from our studies is related to the shapes and granularity
of the tasks of workflows in the ensembles. The benchmarks workflows that we
tested have relatively large number of tasks (50--1000). Montage and CyberShake
workflows have fine-grained tasks of short execution times, in the order of
seconds and minutes. LIGO, Epigenomics and SIPHT have longer tasks: from minutes to hours.
Therefore the dynamic algorithms achieved better results on the fine grained
workflows, since the simple random-based allocation strategy leads to good
load-balancing and high resource utilization. On the other hand, when tasks are
longer, it becomes more beneficial to apply a static planning strategy of SPSS.

Our current study reveals multiple paths for future work. One interesting
question would be to answer how the static and dynamic algorithms perform in the
cases of increased uncertainties and failures, that have important influence
on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
important goal would be to extend our application and infrastructure model to
include various data storage options available on clouds. The experimental study
on that subject~\cite{Juve2010} suggests that data demands of scientific
workflows have a high impact not only on execution time but also on costs on
commercial clouds. Finally, we would like to address the heterogeneity of the
infrastructure, as multiple instance types and providers, including private
and community clouds will make the problem more complex and
challenging~\cite{Marshall2010,vockler11,Juve2010}.

>>>>

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enables the algorithms to finish almost all the workflows. In three of five 
cases the SPSS algorithm performs better for tighter constraints: when the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. For these 
applications the large number of small tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the tested workflows the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

@@

In this paper we addressed the interesting and important new problem of 
scheduling and resource provisioning for scientific workflow ensembles 
on IaaS clouds. The problem is different from previous work on grid and 
utility grid scheduling in that cloud infrastructures provide more control 
over the resources, so the resource provisioning plan can be adjusted 
according to the requirements of the application. Therefore the problem 
space becomes larger: it requires not only an efficient mapping of tasks
to available resources, but also the selection of the best resource 
provisioning plan. Formulating the problem as a maximization of the
number of prioritized workflows completed from the ensemble also 
requires workflows to be admitted or rejected based on their estimated 
resource demands. We believe that this bi-constrained problem is highly 
relevant because such constraints are commonly imposed on many real-world 
projects.

We have considered both static and dynamic scheduling approaches and 
developed the SPDS, DPDS, WA-DPDS and SPSS algorithms based on these 
strategies. The algorithms were evaluated on ensembles of synthetic 
workflows, which were generated based on real scientific applications. For 
the purposes of evaluation we have developed a simulator that models: the 
cloud infrastructure, and a workflow engine with tightly-coupled scheduler 
and provisioner modules. The results of our simulation studies indicate 
that the two algorithms that take into account information about the 
structure of the workflow and task runtime estimates (WA-DPDS and SPSS) 
yield better results than the simple on-line priority-based scheduling 
strategy (DPDS).

During our work, several interesting lessons were learned about running 
workflow ensembles on cloud infrastructures. One result is that the simple 
strategy used by DPDS performs relatively well in terms of number of 
workflows completed and the high resource utilization it can achieve. However,
the extra workflows may be of lower-priority, and the high utilization
may be achieved by running tasks from workflows that cannot be completed
given the constraints. On the other hand, DPDS may be the only choice when 
estimates of task runtimes are not available in advance. We also observed
that the WA-DPDS and SPSS algorithms are able to find better quality 
solutions in terms of the target metrics: this is primarily a result of the 
admission algorithms rejecting workflows that are too large to be completed 
given the constraints, and accepting more smaller workflows that can be 
completed.

Another general observation is that although IaaS clouds provide elasticity
and the illusion of unlimited resources, which makes it possible to provision
more VMs to run workflows in parallel, it turns out that higher resource
utilization and better cost effectiveness can be achieved by limiting 
parallelism as much as possible. This observation comes from the fact that 
starting and stopping VMs incurs overheads which can be amortized by running
each VM for as long as possible. This effect was confirmed in the case of 
the SPSS algorithm for Montage and CyberShake workflows with small budgets
where SPSS opted to provision many VMs in parallel, and shut them down early,
so that tasks early in the ensemble workflows could be completed by their 
deadlines. Similarly, a high degree of parallelization leads to lower 
resource utilization and increased cost. This result, visible in 
Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

Another lesson learned from our studies is the effect of task granularity on 
workflow ensembles. The benchmark workflows that we tested have relatively 
large number of tasks (50--1000), but the Montage and CyberShake workflows have 
fine-grained tasks with execution times on the order of seconds to minutes, 
while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
ranging from several minutes to hours. We observed that the dynamic 
algorithms achieved better results on the fine grained workflows since the 
simple random scheduling strategy leads to good load-balancing and high 
resource utilization for these ensembles. On the other hand, when tasks are
longer, it becomes more beneficial to apply the more sophisticated static 
scheduling strategy of SPSS.

Our current study suggests many areas for future work. One interesting
issue that we plan to investigate is how the static and dynamic algorithms 
perform with uncertain task runtimes and failures, which are known to have 
an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}. 
We also plan to extend our application and infrastructure model to 
include the various data storage options available on clouds. A previous 
experimental study on that subject~\cite{Juve2010} suggests that the data 
demands of scientific workflows have a large impact on not only the 
execution time, but also on the cost of workflows in commercial clouds. 
Finally, we plan to investigate heterogeneous environments that include 
multiple VM types and cloud providers, including private and community clouds, 
which will make the problem even more complex and challenging~\cite{Marshall2010,vockler11,Juve2010}.
